\chapter*{Abstract}

Reoccurring local outbreaks of new, highly pathogenic strains of the \gls{IAV}, picture a unnoticed but still persisting major danger to the whole human population, that reached global extend with millions of deaths several times in the past. Due to the lack of a real cure, resort to vaccines producing varying levels of immunization with yearly expiration is inevitable. For better preparation on possible future pandemics, enlarging the knowledge of the \gls{IAV} is crucial. High evolution-rates by more drastic mutation mechanisms of the \gls{IAV}, with an aged classification containing little insight, complicate accurate novel research though. This thesis, thus, serves the elaboration of a pipeline usable on all existing and subsequently sequenced genomes of \gls{IAV}, to renew the current classification. Instead of being alignment based, this method utilizes the better scalablity of $k$-mer frequency vectors with a the novel hybrid clustering implementation of \texttt{HDBSCAN}, connecting hierarchical with density-based clustering. Most appropriate parameters were thoroughly tested and selected by different validation techniques. For best preservation of the high amount of information included in the vectors used in the clustering, different tools were tested for the best representation in a clusterable dimension. Thereby, a workflow combining a novel vector clustering method with validated parameters, posterior to a dimension reduction, preserving a high amount of information, was proposed. 